---
title: "Final_project_luke"
author: "Luke Awino"
date: '2022-06-14'
output: word_document
---
#Load the packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(corrplot)
library(zoo)
library(Hmisc)
library(e1071)
```
#Load the datasets
```{r}
df1 <- read.csv('/Users/Luke-Workstation/Desktop/ADS 503/ADS-503-Team-3/clean title/anar_primary.csv')
df2 <- read.csv('C:/Users/Luke-Workstation/Desktop/ADS 503/ADS-503-Team-3/clean title/anar_lower_secondary.csv')
df3 <- read.csv('C:/Users/Luke-Workstation/Desktop/ADS 503/ADS-503-Team-3/clean title/anar_upper_secondary.csv')
df4 <- read.csv('C:/Users/Luke-Workstation/Desktop/ADS 503/ADS-503-Team-3/clean title/oos_primary.csv')
df5 <- read.csv('C:/Users/Luke-Workstation/Desktop/ADS 503/ADS-503-Team-3/clean title/oos_lower_secondary.csv')
df6 <- read.csv('C:/Users/Luke-Workstation/Desktop/ADS 503/ADS-503-Team-3/clean title/oos_upper_secondary.csv')
df7 <-read.csv('C:/Users/Luke-Workstation/Desktop/ADS 503/ADS-503-Team-3/clean title/completion_primary.csv')
df8 <- read.csv('C:/Users/Luke-Workstation/Desktop/ADS 503/ADS-503-Team-3/clean title/foundational_reading.csv')
df9 <- read.csv('C:/Users/Luke-Workstation/Desktop/ADS 503/ADS-503-Team-3/clean title/foundational_numeracy.csv')
head(df1)
```
#Column names for upper & lower limits for Total, Children without functional disabilities, and Children with functional disabilities
do not have the correct names. The names of these columns are assigned by following:
```{r}
colnames(df1)[9:10] <- c("Total_Upper_Limit", "Total_Lower_Limit")
colnames(df1)[12:13] <- c("Children_without_functional_difficulties_upper_limit", "Children_without_functional_difficulties_lower_limit")
colnames(df1)[15:16] <- c("Children_with_functional_difficulties_upper_limit", "Children_with_functional_difficulties_lower_limit")
colnames(df2)[9:10] <- c("Total_Upper_Limit", "Total_Lower_Limit")
colnames(df2)[12:13] <- c("Children_without_functional_difficulties_upper_limit", "Children_without_functional_difficulties_lower_limit")
colnames(df2)[15:16] <- c("Children_with_functional_difficulties_upper_limit", "Children_with_functional_difficulties_lower_limit")
colnames(df3)[9:10] <- c("Total_Upper_Limit", "Total_Lower_Limit")
colnames(df3)[12:13] <- c("Children_without_functional_difficulties_upper_limit", "Children_without_functional_difficulties_lower_limit")
colnames(df3)[15:16] <- c("Children_with_functional_difficulties_upper_limit", "Children_with_functional_difficulties_lower_limit")
colnames(df4)[9:10] <- c("Total_Upper_Limit", "Total_Lower_Limit")
colnames(df4)[12:13] <- c("Children_without_functional_difficulties_upper_limit", "Children_without_functional_difficulties_lower_limit")
colnames(df4)[15:16] <- c("Children_with_functional_difficulties_upper_limit", "Children_with_functional_difficulties_lower_limit")
colnames(df5)[9:10] <- c("Total_Upper_Limit", "Total_Lower_Limit")
colnames(df5)[12:13] <- c("Children_without_functional_difficulties_upper_limit", "Children_without_functional_difficulties_lower_limit")
colnames(df5)[15:16] <- c("Children_with_functional_difficulties_upper_limit", "Children_with_functional_difficulties_lower_limit")
colnames(df6)[9:10] <- c("Total_Upper_Limit", "Total_Lower_Limit")
colnames(df6)[12:13] <- c("Children_without_functional_difficulties_upper_limit", "Children_without_functional_difficulties_lower_limit")
colnames(df6)[15:16] <- c("Children_with_functional_difficulties_upper_limit", "Children_with_functional_difficulties_lower_limit")
colnames(df7)[9:10] <- c("Total_Upper_Limit", "Total_Lower_Limit")
colnames(df7)[12:13] <- c("Children_without_functional_difficulties_upper_limit", "Children_without_functional_difficulties_lower_limit")
colnames(df7)[15:16] <- c("Children_with_functional_difficulties_upper_limit", "Children_with_functional_difficulties_lower_limit")
colnames(df8)[9:10] <- c("Total_Upper_Limit", "Total_Lower_Limit")
colnames(df8)[12:13] <- c("Children_without_functional_difficulties_upper_limit", "Children_without_functional_difficulties_lower_limit")
colnames(df8)[15:16] <- c("Children_with_functional_difficulties_upper_limit", "Children_with_functional_difficulties_lower_limit")
colnames(df9)[9:10] <- c("Total_Upper_Limit", "Total_Lower_Limit")
colnames(df9)[12:13] <- c("Children_without_functional_difficulties_upper_limit", "Children_without_functional_difficulties_lower_limit")
colnames(df9)[15:16] <- c("Children_with_functional_difficulties_upper_limit", "Children_with_functional_difficulties_lower_limit")
```
### Removal of the unnecessary subcategories (row 1)
```{r}
df1_clean <- as.data.frame(df1[-1,])
df2_clean <- as.data.frame(df2[-1,])
df3_clean <- as.data.frame(df3[-1,])
df4_clean <- as.data.frame(df4[-1,])
df5_clean <- as.data.frame(df5[-1,])
df6_clean <- as.data.frame(df6[-1,])
df7_clean <- as.data.frame(df7[-1,])
df8_clean <- as.data.frame(df8[-1,])
df9_clean <- as.data.frame(df9[-1,])
```
#After removing the unneeded subcategory row, the columns are combined into one for convenience.
```{r}
df_comb <- cbind(df1_clean[,1:16], df2_clean[,6:16], df3_clean[,6:16],
                 df4_clean[,6:16], df5_clean[,6:16], df6_clean[,6:16],
                 df7_clean[,6:16], df8_clean[,6:16], df9_clean[,6:18])
dim(df_comb)
df_comb
```


```{r}
dfc <- df_comb[,c(5,8,11,14,19,22,25,30,33,36,41,44,47,52,55,58,63,66,69,74,77,80,85,88,91,96,99,102)] # Excluding explanatory and limit values
total <- dfc[c(1,6,11,16,21,26,31,36,41,46,51,56,61,66,71,76,81,86,91,96,
               101,106,111,116,121,126,131,136,141,146,151,156),] # subsetting total values
male <- dfc[c(2,7,12,17,22,27,32,37,42,47,52,57,62,67,72,77,82,87,92,97,
              102,107,112,117,122,127,132,137,142,147,152,157),] # subsetting male values
female <- dfc[c(3,8,13,18,23,28,33,38,43,48,53,58,63,68,73,78,83,88,93,98,
                103,108,113,118,123,128,133,138,143,148,153,158),] # subsetting female values
urban <- dfc[c(4,9,14,19,24,29,34,39,44,49,54,59,64,69,74,79,84,89,94,99,
               104,109,114,119,124,129,134,139,144,149,154,159),] # subsetting urban values
rural <- dfc[c(5,10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90,95,
               100,105,110,115,120,125,130,135,140,145,150,155,160),] # subsetting rural values
```

#The dataset has not been separated into 5 different subsets.

```{r}
colnames(total) <- c("level","primaryANAR_total","primaryANAR_nodiff_total","primaryANAR_diff_total",
                     "lowsecondaryANAR_total","lowsecondaryANAR_nodiff_total","lowsecondaryANAR_diff_total",
                     "uppsecondaryANAR_total","uppsecondaryANAR_nodiff_total","uppsecondaryANAR_diff_total",
                     "primaryOOS_total","primaryOOS_nodiff_total","primaryOOS_diff_total",
                     "lowsecondaryOOS_total","lowsecondaryOOS_nodiff_total","lowsecondaryOOS_diff_total",
                     "uppsecondaryOOS_total","uppsecondaryOOS_nodiff_total","uppsecondaryOOS_diff_total",
                     "primarycomp_total","primarycomp_nodiff_total","primarycomp_diff_total",
                     "reading_total","reading_nodiff_total","reading_diff_total",
                     "numeric_total","numeric_nodiff_total","numeric_diff_total")
colnames(male) <- c("level","primaryANAR_male","primaryANAR_nodiff_male","primaryANAR_diff_male",
                    "lowsecondaryANAR_male","lowsecondaryANAR_nodiff_male","lowsecondaryANAR_diff_male",
                    "uppsecondaryANAR_male","uppsecondaryANAR_nodiff_male","uppsecondaryANAR_diff_male",
                    "primaryOOS_male","primaryOOS_nodiff_male","primaryOOS_diff_male",
                    "lowsecondaryOOS_male","lowsecondaryOOS_nodiff_male","lowsecondaryOOS_diff_male",
                    "uppsecondaryOOS_male","uppsecondaryOOS_nodiff_male","uppsecondaryOOS_diff_male",
                    "primarycomp_male","primarycomp_nodiff_male","primarycomp_diff_male",
                    "reading_male","reading_nodiff_male","reading_diff_male",
                    "numeric_male","numeric_nodiff_male","numeric_diff_male")
colnames(female) <- c("level","primaryANAR_female","primaryANAR_nodiff_female","primaryANAR_diff_female",
                      "lowsecondaryANAR_female","lowsecondaryANAR_nodiff_female","lowsecondaryANAR_diff_female",
                      "uppsecondaryANAR_female","uppsecondaryANAR_nodiff_female","uppsecondaryANAR_diff_female",
                      "primaryOOS_female","primaryOOS_nodiff_female","primaryOOS_diff_female",
                      "lowsecondaryOOS_female","lowsecondaryOOS_nodiff_female","lowsecondaryOOS_diff_female",
                      "uppsecondaryOOS_female","uppsecondaryOOS_nodiff_female","uppsecondaryOOS_diff_female",
                      "primarycomp_female","primarycomp_nodiff_female","primarycomp_diff_female",
                      "reading_female","reading_nodiff_female","reading_diff_female",
                      "numeric_female","numeric_nodiff_female","numeric_diff_female")
colnames(urban) <- c("level","primaryANAR_urban","primaryANAR_nodiff_urban","primaryANAR_diff_urban",
                     "lowsecondaryANAR_urban","lowsecondaryANAR_nodiff_urban","lowsecondaryANAR_diff_urban",
                     "uppsecondaryANAR_urban","uppsecondaryANAR_nodiff_urban","uppsecondaryANAR_diff_urban",
                     "primaryOOS_urban","primaryOOS_nodiff_urban","primaryOOS_diff_urban",
                     "lowsecondaryOOS_urban","lowsecondaryOOS_nodiff_urban","lowsecondaryOOS_diff_urban",
                     "uppsecondaryOOS_urban","uppsecondaryOOS_nodiff_urban","uppsecondaryOOS_diff_urban",
                     "primarycomp_urban","primarycomp_nodiff_urban","primarycomp_diff_urban",
                     "reading_urban","reading_nodiff_urban","reading_diff_urban",
                     "numeric_urban","numeric_nodiff_urban","numeric_diff_urban")
colnames(rural) <- c("level","primaryANAR_rural","primaryANAR_nodiff_rural","primaryANAR_diff_rural",
                     "lowsecondaryANAR_rural","lowsecondaryANAR_nodiff_rural","lowsecondaryANAR_diff_rural",
                     "uppsecondaryANAR_rural","uppsecondaryANAR_nodiff_rural","uppsecondaryANAR_diff_rural",
                     "primaryOOS_rural","primaryOOS_nodiff_rural","primaryOOS_diff_rural",
                     "lowsecondaryOOS_rural","lowsecondaryOOS_nodiff_rural","lowsecondaryOOS_diff_rural",
                     "uppsecondaryOOS_rural","uppsecondaryOOS_nodiff_rural","uppsecondaryOOS_diff_rural",
                     "primarycomp_rural","primarycomp_nodiff_rural","primarycomp_diff_rural",
                     "reading_rural","reading_nodiff_rural","reading_diff_rural",
                     "numeric_rural","numeric_nodiff_rural","numeric_diff_rural")
df_c <- cbind(total,male[,2:28],female[,2:28],urban[,2:28],rural[,2:28])
head(df_c)
dim(df_c)
#fix(df_c)
```
#Now, the columns have been re-named.
```{r filling in NA and Not Specified values in level column}
df_c[,1]
# Among the 32 values, one NA value and on Not Specified (26th row) value were noted.
df_c[,1] <- na.fill(df_c[,1], "Less Developed")
df_c[26,1] <- "More Developed"
summary(df_c)
#fix(df_c)
```
#The levels have now been filled in, but the values are all in characters, not numeric.
```{r changing the numeric values to appropriate dataset}
df_c[,2:136] <- lapply(df_c[,2:136], FUN = function(y){as.numeric(y)})
summary(df_c)
```
#plot the misssing values

```{r}
hist(df_c)

```


# check for skewness
```{r}
skewness(df_c[,2:136], na.rm = FALSE)
```
#check for correlation
```{r}
corrplot::corrplot(cor(df_c[,2:136]))
```
there are too many predictors 

#The values are now all numeric, except the levels column, and there are NA values in these numeric variables.
```{r filling in NA numeric values & excluding highly correlated variables - Pre-processing}

df_imp <- preProcess(df_c[,2:136],method =  "knnImpute", k=1)
df_imputed <- predict(df_imp,df_c)
summary(df_imputed) # Now, the NA are filled in with values computed with KNN.
tooHigh <- findCorrelation(cor(df_imputed[,2:136]), 0.8)
df_filtered <- df_imputed[,-tooHigh]
level_final <- as.factor(make.names(df_c[,1]))
dim(df_filtered)
summary(level_final)
df_final <- cbind(level_final, df_filtered)
head(df_final)
#fix(df_final)
```
# check for skewness in the dataset
```{r}
skewness(df_final)
```
#plot the correlation for final dataset

```{r}
corrplot::corrplot(cor(df_final[,2:6]))
```
#
```{r}
table(df_final$level_final)
```
the dataset has 13 observations and most of the the development regions are least developed , less developed both having 28 combined observations, more developed with 3 and 1 unidentified.

```{r data split & Naive Bayes model}
df_split <- sort(sample(nrow(df_final), nrow(df_final)*0.8))
df_train <- df_final[df_split,]
df_test <- df_final[-df_split,]
library(naivebayes)
df_nb <- naive_bayes(df_train[,2:6], df_train[,1])
df_pred <- predict(df_nb, df_test[,2:6])
#confusionMatrix(df_pred,df_test[,1])
```

# Random forest model
```{r}
mtryValues <- seq(1,10,1)

#
ctrl <- trainControl(method = "cv",
                     summaryFunction = multiClassSummary,
                     classProbs = TRUE,
                     savePredictions = TRUE)
#fix(df_final)
dim(df_final)
library(randomForest)
set.seed(326) 

rfFit <- train(x = df_final[,2:13], 
                y = df_final$level_final,
                method = "rf",
                ntree = 1000,
                tuneGrid = data.frame(mtry = mtryValues),
                metric = "ROC",
                trControl = ctrl)
rfFit

?make.names 
```



```{r}
library(gbm)

```

